Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Commenting out logisticPCA stuff for now.

In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures?; 2) what functions are leaders most commonly serving across cultures?; 3) what are the most salient costs and benefits of leadership for both leaders and followers?; and 4) how does variation across these measures relate to population-level variation?

@@

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record.

Our second goal was to discover "features" in our data, e.g., to determine if particular variables tended to covary or cluster together. Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001]. Or they might suggest new and theoretically important dimensions of leadership.

There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA].

Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors.

@@

We also performed hierarchical cluster analysis on our quality and function variables. We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record.

Our fourth goal was to explore associations between the principle components of leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size. We used linear mixed effects models from the lme4 package [@bates2015], with PC1 and PC2 as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.

>>>>

In sum, we aim to empirically and systematically answer the following questions: 1) what are the most frequent qualities of leaders across cultures? 2) What functions are leaders most commonly serving across cultures?; 3) What "features" exist in our data, e.g., do particular variables tended to covary or cluster together? Such features might lend support to existing theories; for instance they might correspond to "prestige" and "dominance" [@henrich_evolution_2001]. Or they might suggest new and theoretically important dimensions of leadership. 4) What are the most salient costs and benefits of leadership for both leaders and followers? And 5) how does variation across these measures relate to population-level variation?

@@

We analyzed data from our researcher-coded operationalized variables at the text-record level and at the culture-level. Our first goal was to assess the proportion of text records and cultures that provided evidence for each of our coded variables to discover the qualities, functions, benefits and costs of leadership that are most and least frequently noted in the ethnographic record. We computed confidence intervals for the proportion of text records supporting each variable using generalized linear mixed effects models and random intercepts based on document authors nested within cultures. We computed confidence intervals for the proportion of cultures with evidence for each variable using a cluster bootstrap.

We then explored "features" in our data. There are numerous approaches to detecting features in matrix data such as ours, many of which employ various dimensionality reduction and clustering techniques. We used hierarchical cluster analysis on our quality and function variables. We computed distance between our binary vectors using both $1-cor(x, y)$ and asymmetric binary distance (Jaccard distance), and agglomerated using the Ward algorithm. We assessed the robustness of the clusters using the pvclust package [@suzuki_pvclust2006], which computes an unbiased probability that a cluster appears in bootstrapped samples. We used 10,000 bootstrapped samples.

<!-- To reduce the dimensionality of the `r length(quality_vars)` leader quality variables and the `r length(function_vars)` leader function variables, all of which take binary values (0, 1), we used logistic PCA from the logisticPCA package. This algorithm projects the natural parameters of the saturated model onto the lower $k-$dimension space that minimizes the Bernoulli deviance [@logisticPCA]. Prior to analysis we removed all rows that had no evidence for any variable, which resulted in `r nrow(pca_data_qualities)` text records for the leader quality variables and `r nrow(pca_data_functions)` text records for the leader function variables. We used cross-validation to optimize rank, $k$, and a tuning parameter, $m$ (the value to approximate the saturated model), per recommendation of the package authors. -->

@@

Our third goal was to determine which leader and follower benefits and costs were most commonly mentioned in the ethnographic record, computing 95% CIs as above.

Our fourth goal was to explore associations between the most common leader qualities and functions, and culture-level variables such as mode of subsistence, social complexity, population density, and group size. We used linear mixed effects models from the lme4 package [@bates2015], with common qualities and functions as outcome variables, culture-level variables as predictors, and documents nested within cultures as grouping variables for random intercepts.
